### 2024-02-15 revision/reply_to_reviewer2.tex

Thanks for the careful reading and for your helpful comments and
suggestions.  Please find below point-by-point replies (in black) to
your comments and questions (which are reprinted in blue). To give you
an overview of all the changes in the paper, we also provide a
diff-document that highlights the changes between the initial
submission and this re-submission.\\[1ex]

\referee{I have read the paper with pleasure and have the following
  suggestions for improvement:\\}

\begin{enumerate}
\item \referee{Interesting work, generally well-written. Perhaps a
  small illustrative example with a few figures at the beginning could
  help understand the procedure intuitively.}

  \noindent{Frog example with a=1.
  	Structure: local non-negative kernel
(0) Matrix itself column highlighted, show impulse response (kernel\_impulse\_illustration.pdf)
(1) moments. Show one impulse with ellipsoid with widths illustrated
(2) impulse response batches
(3) RBF interpolation (?), evaluation of entries. Columns we know, and entry we don't know. 
(4) H-matrix (structure only)
}

@@

\item \referee{possibly interesting references for estimating Hessians
  in linearised seismic imaging include:
  \url{https://doi.org/10.1016/j.acha.2007.06.007},
  \url{https://doi.org/10.1190/1.2836323}}

  \noindent{We added text in the paragraph at the end of Section 2 to discuss the (lack of) applicability of the method to wave inverse problems and added these references.}

\item \referee{How about comparison with other methods, like those
  proposed in [7],[52]? and you say a bit more? it is hard to conclude
  that your method will always outperform these methods.}

    \noindent{For ref 7: Ask Tucker if we could do HODLR on the
      Hessian coming from ice mountain problem and record the number
      of mat-vecs needed to achieve a certain accuracy, something like
      Figure 8 in teh HODLR paper (DONE).

      Nick will do some estimates to
      compare things with ref 52.}

### 2024-02-15 paper_sisc_first_revision/localpsf.tex
fixed sqrt bug when calculating GLR FRO rank
include rsvd data in paper
rsvd data seems correct
wrote up [52] h matrix discussion / back of envelope calculation

Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}. 
These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,Hartland_2023}. 
Although these methods are promising,
the required number of operator applications is still large (e.g., hundreds to thousands).
In this paper, we also form an H-matrix Hessian approximation. However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques. By using this two stage approach, we reduce the number of operator applications to a few dozen at most.

@@

\begin{table}
	\begin{center}
		{\small
		\begingroup
		\setlength{\tabcolsep}{8pt}
		\renewcommand{\arraystretch}{1.0}
		\begin{tabular}{c| c | c c c c c}
						& Error tol.    &	\#applies PSF   & \#applies HODLR 	& rank \\ 
			\hline 
						& 20\%    		& 11  				& 592				& 37  \\ 
			$L=1$ 		& 10\% 			& 15  				& 772				& 67  \\ 
						& 5\% 			& 23   				& 924				& 100 \\ 
			\hline
						& 20\%    		& 8  				& 852 				& 143 \\ 
			$L=1/2$ 	& 10\%    		& 10  				& 1144				& 256 \\ 
						& 5\% 	  		& 12   				& 1404				& 377 \\ 
			\hline
						& 20\%    	 	& 7					& 932				& 294  \\ 
			$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 525  \\ 
						& 5\% 			& 8   				& 1520 				& 767  \\ 
		\end{tabular}
		\endgroup
	}
	\end{center}
	\caption{(Spatially variant blur) Comparison of cost to approximate the blur kernel from Equation \eqref{eq:frog_kernel} using the PSF method, the randomized HODLR (hierarchical off diagonal low rank) method, and GLR (global low rank) approximation using randomized SVD. The quantity $L$ scales the width of the impulse responses, hence it influences the rank of the operator. Large $L$ means low rank, and small $L$ means high rank. The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$. The remaining three columns show the number of operator applies require to achieve the given error tolerance, using the PSF, HODLR, and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}

>>>>

Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}. 
These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,Hartland_2023}. 
Although these methods are promising,
the required number of operator applications is still large (e.g., hundreds to thousands). For example, using the method in \cite{LinYing11}, the required number of operator applies to construct an $H^1$ matrix with hierarchical rank $r$ for problems in a 2D domain discretized with a regular grid is $\#\text{levels} \cdot 64 \cdot (r + c)$, where $\#\text{levels}$ is the depth of the hierarchical partitioning, $r$ is the rank of the blocks (hierarchical rank), and $c$ is an oversampling parameter (See \cite[Section 2.4]{LinYing11}). On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, the works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies. In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of blur operators with modest (e.g., $10\%$) relative error.
%while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
In this paper, we also form an H-matrix approximation. However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques. By using this two stage approach, we reduce the number of operator applications to a few dozen at most.

@@

\begin{table}
	\begin{center}
	{\small
	\begingroup
	\setlength{\tabcolsep}{8pt}
	\renewcommand{\arraystretch}{1.0}
	\begin{tabular}{c| c | c c c c c}
				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
	\hline 
				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
	$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
	\hline
				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
	$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
	\hline
				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
	$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
	\end{tabular}
	\endgroup
	}
	\end{center}
	\caption{(Spatially variant blur) Comparison of cost to approximate the blur kernel from Equation \eqref{eq:frog_kernel} using the PSF method, the randomized HODLR (hierarchical off diagonal low rank) method, and GLR (global low rank) approximation using randomized SVD. The quantity $L$ scales the width of the impulse responses, hence it influences the rank of the operator. Large $L$ means low rank, and small $L$ means high rank. The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$. The remaining three columns show the number of operator applies require to achieve the given error tolerance, using the PSF, HODLR, and GLR methods.}
	\label{tab:frog_psf_vs_hodlr_vs_glr}
\end{table}
